---
title: "Visualizations with ggplot"
Author: "Juliet Thistleton"
output: html_notebook
---

Using the ggplot package, We will explore a data set that keeps track of the name of the video game released, the platofrom it was released onm the release date, a summary of the game, and the User review rating of the game. 

This data set was taken from https://www.kaggle.com/datasets/maso0dahmed/video-games-data.

# Required Packages 
```{r}
library(tidyverse) 
library(janitor)
library(here)
library(lubridate)
```
# Defining helpful variables
```{r}
rotate = theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust= 1))
```



# Import the data 
```{r}
df1 = read.csv(here("data_raw", "video_games.csv")) |> clean_names()
```


# Question 1: Which video game console had the most games released for it beteween April 1, 2001, and Sepetember 9, 2021 according to this data set?

```{r}
df2 = df1 |> 
  select(-summary) |>
  group_by(platform) |>
  count()

```

## Making a visualization to see the console with the most releases
```{r}
ggplot(df2, (aes(x = platform, y = n))) + geom_col() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust= 1))
```
From this graph, we can see that the PC has had the most releases of all the platform included in this data set. 

# Question 2: What is the average review score for each console's games released?

## Coercing user_revew from character to int
```{r}
df3 = df1 |>
  group_by(user_review) |>
  count()


ggplot(df3, aes(user_review, n)) + geom_col() + theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust= 1))
```


```{r}
aggregate(df3$user_review, list(df3$platform), FUN = mean)
```


